---
title: "Project 1-JSON Vinette"
author: "Grant Swigart"
date: "6/7/2020"
output: 
  rmarkdown::github_document:
        toc: yes
        toc_depth: 3
---



# JSON
## What is it, where does it get used, and why is it a good way to store data? 

JSON stands for Javascript Object Notation. It is a file type for storing and transferring data. It was originally invented in the early 2000's to help servers communicate with browsers. 


Lets look at some examples of a JSON objects!

This is what a JSON object typically looks like. Typically you can access each part of the JSON data by calling the name of the larger group, similair to a list in R. However, for us to work with this data in R we need to convert this  data in a usable format. 

```{r JSON intro,warning=FALSE,message=FALSE}
library(jsonlite)
hockey_list <- list(Wayne = list(city = "Edmonton",last_name = "Gretzky"), 
          Sidney= list(city = "Pittsburgh",last_name = "Crosby"),
          Gordie= list(city = "Detroit",last_name = "Howe"))
json<-toJSON(hockey_list, pretty = TRUE, auto_unbox = TRUE)
print(json)
```

## What package should I use to read JSON data into R?

There ae three main packages that read JSON data into R. 

* rjson
* RJSONIO
* jsonlite

One of the major benefits of jsonlite is that it better maps JSON data into R into data types that are more heavely used such as dataframes, lists, and matrices Also, it is better in maintaining the way that missing values ae coded. It also provides more information than the other packages when an error occurs. JSON lite is also effecient, especially when compared to rjson.

```{r JSON library}
json_data<-fromJSON(json,flatten=TRUE)
print(json_data)
```

Notice that jsonlite correctly maps the data to the original data structure (list of 3 lists.). Now that we have this data in a usable data structure we can use other tools to summarize and analyze the data. 


## What the puck. Lets look at some JSON data!

Here are the packages that we will use to analyze our data.
```{r setup,message=FALSE,warning=FALSE}
library(devtools)
library(tidyverse)
library(httr)
library(knitr)
```


We will be looking the data available from the [NHL records API](https://gitlab.com/dword4/nhlapi/-/blob/master/records-api.md).An API is a application program interface and is essentially software that allows computers to communicate with one another. Think of how your browser (chrome, edge, firefox) follows a url to access information from a webpage. That webpage sends us HTML code that your computer breaks down into an interface. In this case we want to use R to follow a URL to request data from the NHL records server. The text data returned from this survey is JSON structured. Using the above package we can then convert this data in a R dataframe, list, or tibble which we can analyze.

API's usually have several paramaters that you can specify to access different information. This API has several tables that you can access tha have information on player records, team records, attendance, and draft information. We can also filter some of this information according to a specific team id we want to analyze.

Lets get aquanted with using an API. An API hase a base url that is used for all data requests and then has parameters we want to set. In this case we want to look at the New Jersey Devils which has a teamid of 1. Additiaonlly we want to look at the skater records for this franchise. After combining the text for this text, we use a the GET function to submit our request. Printing this request we are able to see a status code. A status coded of 200 means we requested data from a URL that exists. If you receive a 404 this means the URL was not found. 

```{r example request}
base_url <- 'https://records.nhl.com/site/api'
id<-'1'
build_url=paste0(base_url,'/franchise-skater-records?cayenneExp=franchiseId=',id)
get_request<-GET(build_url)
print(get_request)
```


Now that we have this request we want to turn the request object into text. We do this by using the content function and specifying the output and encoding. Then we use fromJSON to convert the data into a usable object. The flatten option turns nested data frames into a single data frame. 

```{r example import}
request_text<-content(get_request,"text",encoding='UTF-8')
request_list<-fromJSON(request_text,flatten=TRUE)
request_data<-request_list$data
print(head(request_data,5))
```

Now that we understand how APIs work, let's create a larger function to request multiple different tables from the API. We need to include error checks to make sure the user understands any request mistakes. We also use simplier table names to make the requesting data easier.

```{r API function}
get_nhl<-function(table='franchise',id=''){
  #Convert the table names and franchise id's into string to prevent data issues with logical checks. 
  table<-toString(table)
  id<-toString(id)
  #This list of available ids is taken from the franchise table. 
  available_id<-c("8","41","45","37","10","6","43","51","39","3","16","17","49","26","25","4","5","19","7","23","20","2","1","15","22","12","21","53","28","9","14","24","13","18","52","29","30","54")
  
  #Basic logical checks to see if there have been any incorrect API calls or combinations of ids with the wrong table. 
  if (id!='' & (!table %in% c('skater-records','goalie-records','season-records'))){
                stop('ERROR: Franchise Id is only used for these tables:\n
                     "skater-records","goalie-records","season-records"')
  }
  else if (id!='' & (!id %in% available_id)){
                stop(paste("ERROR: The available id's are ",paste0(available_id,collapse=" ")))
  }

  
  #This is the base URL for the API we want to access
  base_url <- 'https://records.nhl.com/site/api'
  #We then build a different url based upon the data table specified.
  if(table=='franchise'){
    build_url=paste0(base_url,'/franchise')
  }
  else if(table=='team-totals'){
    build_url=paste0(base_url,'/franchise-team-totals')
  }
  else if(table=='player'){
    build_url=paste0(base_url,'/player')
  }
  else if(table=='draft'){
    build_url=paste0(base_url,'/draft')
  }
  else if(table=='attendance'){
    build_url=paste0(base_url,'/attendance')
  }
  
  else if(table=='season-records' & id==''){
    build_url=paste0(base_url,'/franchise-season-records')
  }
  else if(table=='goalie-records' & id==''){
    build_url=paste0(base_url,'/franchise-goalie-records')
  }
  else if(table=='skater-records' & id==''){
    build_url=paste0(base_url,'/franchise-skater-records')
  }
  
  else if(table=='season-records' & id!=''){
    build_url=paste0(base_url,'/franchise-season-records?cayenneExp=franchiseId=',id)
  }
  else if(table=='goalie-records' & id!=''){
    build_url=paste0(base_url,'/franchise-goalie-records?cayenneExp=franchiseId=',id)
  }
  else if(table=='skater-records' & id!=''){
    build_url=paste0(base_url,'/franchise-skater-records?cayenneExp=franchiseId=',id)
  }
  
  else {
    stop('ERROR: The available tables are "franchise","team-totals", "season-records", "goalie-records", "skater-records", "player","draft","attendance"')
  }
  get_request<-GET(build_url)
  #Stop if the url request wa not to an acceptable UR?L. 
  if (get_request$status_code==404){
    stop('Status Code 404.\n Inccorect URL. This is likely due to askinig for an unavailable table.\n
        These are tha available tables. Use the "franchise" table to find the list of franchise ids.
         "team-totals","season-records","goalie-records","skater-records","player","draft","attendance"')
  }
  request_text<-content(get_request,"text",encoding='UTF-8')
  request_list<-fromJSON(request_text,flatten=TRUE)
  request_data<-request_list$data
  return(request_data)
}
```

Let's request all the data from all the tables the above function can request. We dont have to specify a table for the franchise table because this is the default. 

```{r table}
franchise<-get_nhl()
team_totals<-get_nhl(table='team-totals')
draft<-get_nhl(table='draft')
attendance<-get_nhl(table='attendance')
player<-get_nhl(table='player')
rec_season<-get_nhl('season-records')
rec_goalie<-get_nhl('goalie-records')
rec_skater<-get_nhl('skater-records')
```

Not we want to request the


# Grahs and Visualizations

## Top 10 

Now that we have looked at our data lets view some graphs to identify trends. The rec_skater dataframe may has multiple records for the same player. Lets calcualte some summary statistics for the total points, assists, penalty minutes, games played, and other information. 
```{r}
summary_points<-rec_skater %>%
  mutate(full_name=paste(firstName,lastName)) %>%
  group_by(playerId,full_name,activePlayer,positionCode) %>%
  summarise(seasons=sum(seasons),
            total_points=sum(points),
            games_played=sum(gamesPlayed),
            rookiePoints=sum(rookiePoints),
            penaltyMinutes=sum(penaltyMinutes),
            total_assists=sum(assists),
            total_goals=sum(goals)) %>%
  mutate(ppg=round(total_points/games_played,2),
         ppg_label=paste(as.character(ppg),'ppg')) %>%
  arrange(desc(total_points))
```

Lets look at the players who have the most points scored(goals+assists) of all times. Also lets look at effeciency by adding text that has the points per game for each player .
```{r top 10 points}
ggplot(summary_points %>% head(10),aes(x=reorder(full_name,total_points),y=total_points))+
  geom_col()+ 
  coord_flip()+
  geom_text(aes(label=ppg_label), position=position_dodge(width=0.9), hjust=1.2,vjust=.2)+
  ylab('Points Scored')+
  xlab('Player')+
  ggtitle('Top 10 Points Scored All Time')
```

## Summary Statistics by Position

Next lets break some of the information accross position. Lets group out data by position code and then take the top 10 highest scorers of each position. 

```{r position points table}
top_10_pos<-summary_points %>%
  group_by(positionCode) %>%
  top_n(10,wt=total_points)%>%
  select(positionCode,full_name,total_points,ppg,total_assists,total_goals) 

kable(top_10_pos%>% filter(positionCode=='L') %>% select(-positionCode),caption ="Top 10 Left Wing Scorers" )
kable(top_10_pos%>% filter(positionCode=='R') %>% select(-positionCode),caption ="Top 10 Right Wing Scorers" )
kable(top_10_pos%>% filter(positionCode=='C') %>% select(-positionCode),caption ="Top 10 Center Scorers" )
kable(top_10_pos%>% filter(positionCode=='D') %>% select(-positionCode),caption ="Top 10 Defensive Scorers" )

```

## Performance by Draft Pick

Next lets see what the average ppg for each draft pick. We merged the draft dataframe with our summary statistcs and then calculate our average ppf. We expect to see less ppg for defensemen so we should remove them from our analysis.

```{r PPG by Pick calculations}
points_draft<-summary_points %>% 
  inner_join(draft,by = "playerId") %>%
  filter(positionCode %in% c('L','R','C'),
         seasons>1) %>%
  group_by(overallPickNumber,positionCode) %>%
  summarise(ppg=mean(ppg),N=n())
```


We create a bar chart to view the ppg for each pick and each player. Notably to identify the trend lines lets add the geom_smooth 

```{r PPG by Pick}
ggplot(points_draft,aes(x=overallPickNumber,y=ppg))+
  facet_wrap(~positionCode)+
  geom_bar(stat='identity')+
  geom_smooth(method='loess')+
  ggtitle('Average PPG by Overall Pick Number and Position',subtitle='Seasons>1, Nondefensemen') +
  xlab('Pick number')+
  ylab('Points Per Game')

```

How many people are drafted from each differnt country? We do this by using the table command on the county code variable to count the number of players. Then we use the kable funciton to make it more pretty. 

## Number of Players by Country

```{r contingency table}
table(draft$countryCode) %>%kable()
```

We can also see how the number of players has changed from year to year by graphing the number of usa born and foreign born players over time. I am a NBA fan so I wondered if similarly we would see an increase of foreign born plavers over time. However, it seems clear that foreign born players are drafted more frequentyly. Additionally, the number of foreign born players looks to be more correlated with the total number of draft picks than is the number of us born players. 

```{r drafted by country}
draft_group<-draft %>%
  mutate(foreign=ifelse(countryCode=="USA",'usa','foreign'))%>%
  group_by(foreign,draftYear) %>%
  summarise(N=n()) %>%
  filter(!is.na(foreign)) 

draft_total<-draft %>%
              group_by(draftYear) %>%
              summarise(N=n(),foreign="total")

draft_country<-bind_rows(draft_group,draft_total)



ggplot(draft_country,aes(x=draftYear,color=foreign))+
  geom_line(aes(y=N,group=foreign))+
  ggtitle('Number of USA Players Drafted by Year')+
  xlab('Draft Year')+
  xlab('Nunmber of Players Drafted')


```

Are players from any one nation better? We then compae the distribution of pgg using boxplots. The three distibutions look largely similair. There appear to be more canadian players with a high points per game compared to sweden or the united states. However, it should be noted that there are many more canadian players than sweden or u.s. players. 
```{r PPG vs Country}
country<-summary_points %>% 
  inner_join(draft,by = "playerId") %>%
  filter(countryCode %in% c('CZE','FIN','RUS','SWE','USA','CAN'))


ggplot(country %>% filter(seasons>1,countryCode %in% c('SWE','USA','CAN')),aes(y=ppg))+
  facet_wrap(~countryCode)+
  geom_boxplot()

```

Lastly, lets look at the number of goals and assists by country of origin. Also, we can include the presence in the hockey hall of fame by using a third dimension that shows through the shape of the point.  

```{r CareerPoints}
country_player<-country%>%
  inner_join(player,by=c("birthDate","firstName","height", "lastName"))

ggplot(country_player %>% filter(seasons>1),aes(x=total_assists,y=total_goals,color=countryCode,shape=inHockeyHof))+
  geom_point()+
  geom_smooth(method='lm',se=FALSE,aes(group=countryCode))+
  ggtitle('Career Points and Assists by Country and Presence in Hall of Fame')+
  xlab('Career Assists')+
  ylab('Career Points')

```



